---
title: "Display elements article on terrestrial radiation map"
author: "Folly Christophe"
date: "13 5 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages, results=F, warning=FALSE, message=FALSE, include=FALSE}
library(INLA)
library(Metrics)
library(fields)
library(tidyverse)
library(ggplot2)
library(scoringRules)
library(knitr)
library(gstat)
library(sp)
library(MatrixModels)
library(tmap)
library(tmaptools)
library(viridis)
library(sf)
library(stars)
```

## Introduction

The purpose of this script is to compute the display elements for use in the article on the terrestrial radiation.

Some can be directly exported, some will need modification in PowerPoint or directly in LaTeX.

## Numbers

### Number of measurements

Number of measurements in total `r nrow(readRDS("../Data/processed/aero_data.rds"))`. 

After cleaning, we have `r nrow(readRDS("../Data/processed/aero_data_clean.rds"))`.

Every 15th, we include `r nrow(readRDS("../_15_noflight/Input/aero_cluster.rds"))`.

### Global statistics

The mean after data cleaning is `r mean(readRDS("../Data/processed/aero_data_clean.rds")$TERR)`.

The median after data cleaning is `r median(readRDS("../Data/processed/aero_data_clean.rds")$TERR)`.


## Figures

### Flight path

```{r load_data}
aero_15 <- readRDS("../_15_noflight/Input/aero_cluster.rds")
aero_cleaned <- readRDS("../Data/processed/aero_data_clean.rds")
library(OpenStreetMap)
CH_canton <- as(st_read("C:/Users/folly/Desktop/Homeoffice/Geo_Political/gg2019/ggg_2019-LV03/shp/k4l19.shp"),"Spatial")
CH_lakes <- as(st_read("C:/Users/folly/Desktop/Homeoffice/Geo_Political/gg2019/ggg_2019-LV03/shp/k4s19.shp"),"Spatial")


# buffer Mühleberg
df <- c("Mühleberg KKW", 587100, 202000) #over cheminee
df <- rbind(df, c("Leibstadt KKW", 655850, 272350)) #over cheminee
df <- as.data.frame(df)
names(df) <- c("Site", "X","Y")
df$X <- as.numeric(as.character(df$X))
df$Y <- as.numeric(as.character(df$Y))
coordinates(df) <- ~X+Y
proj4string(df) <- proj4string(CH_canton)
exc_buffer <- st_buffer(st_as_sf(df), 500)

#flight path
exampleflight <- aero_cleaned[aero_cleaned$Flight %in% c("KKM_rekonstRe-eval"),]
temp_bg <- read_osm(exampleflight, ext=1.02)
tm_shape(temp_bg) + tm_rgb() + 
  tm_shape(exampleflight) + tm_dots(col = "red", legend.show = F) +
  tm_layout(frame = F, bg.color = "transparent") + 
  tm_shape(exc_buffer[1,])+tm_borders(col = "blue")
flightpath <- tm_shape(temp_bg) + tm_rgb() + 
  tm_shape(exampleflight) + tm_dots(col = "red", legend.show = F) +
  tm_layout(frame = F, bg.color = "transparent")+ 
  tm_shape(exc_buffer[1,])+tm_borders(col = "blue")
tmap_save(flightpath, "flightpath.jpg")

```


### Coverage with flights and spCV folds

```{r plot_spCV}
spCV <- readRDS("../_15_noflight/spCV_object.rds")
#rescale coords
aero_15$X <- aero_15$X*1000
aero_15$Y <- aero_15$Y*1000
coordinates(aero_15) <- ~X+Y
proj4string(aero_15) <- proj4string(CH_canton)
# get box of example flight

temp_box <- st_polygon(list(as.matrix(rbind(c(st_bbox(temp_bg)[1],st_bbox(temp_bg)[2]),
                 c(st_bbox(temp_bg)[1],st_bbox(temp_bg)[4]),
                 c(st_bbox(temp_bg)[3],st_bbox(temp_bg)[4]),
                 c(st_bbox(temp_bg)[3],st_bbox(temp_bg)[2]),
                 c(st_bbox(temp_bg)[1],st_bbox(temp_bg)[2])))))
temp_box <- st_sf(st_sfc(temp_box), crs = st_crs(temp_bg))
temp_box$text <- "Figure 2" 

#plot
tm_shape(aero_15) + 
  tm_dots(col = "fold",legend.show = FALSE,palette = viridisLite::viridis(4))+
  tm_layout(frame=F,bg.color = "transparent", legend.text.size = 1.2, legend.title.size = 1.4)+
  tm_add_legend(type = "fill", col = viridisLite::viridis(4), labels = c("1","2","3","4"), title = "Fold") + 
  tm_shape(spCV$blocks) + tm_borders(col = "red", alpha = 0.2) + 
  tm_shape(CH_canton) + tm_borders(col = "slategrey") +
  tm_shape(CH_lakes) + tm_fill(col = "cadetblue1",alpha = 0.4) +
  tm_shape(temp_box) + tm_borders(col = "red",lwd = 2) + 
  tm_text(text = "text", col ="red",xmod = 2.5,ymod = -1.1,size=1.4)

# save
displ_folds <- tm_shape(aero_15) + 
  tm_dots(col = "fold",legend.show = FALSE,palette = viridisLite::viridis(4))+
  tm_layout(frame=F,bg.color = "transparent", legend.text.size = 1.2, legend.title.size = 1.4)+
  tm_add_legend(type = "fill", col = viridisLite::viridis(4), labels = c("1","2","3","4"), title = "Fold") + 
  tm_shape(spCV$blocks) + tm_borders(col = "red", alpha = 0.2) + 
  tm_shape(CH_canton) + tm_borders(col = "slategrey") +
  tm_shape(CH_lakes) + tm_fill(col = "cadetblue1",alpha = 0.4)+
  tm_shape(temp_box) + tm_borders(col = "red",lwd = 2) + 
  tm_text(text = "text", col ="red",xmod = 2.5,ymod = -1.1,size=1.4)
tmap_save(tm = displ_folds,filename = "folds_cv.png")

```

### Maps mixed-effects and extended mixed-effects

```{r plot_predictions}
grid_lm <- readRDS("../_15_noflight/Model_LM_inla/Output/grid_results.rds")
grid_me <- readRDS("../_15_noflight/Model_standard/Output/grid_results.rds")
grid_ex <- readRDS("../_15_noflight/Model_extended_priors/Main/Output/grid_results.rds")
grid_sp <- readRDS("../_15_noflight/Model_SPDE/Output/grid_results.rds")

# use quilt plot - save plots manually
quilt.plot(grid_lm$X,grid_lm$Y, grid_lm$mean,nx=300,ny=200,asp=1, nlevel=80, main = "Mean of Prediction",axes=FALSE)
quilt.plot(grid_me$X,grid_me$Y, grid_me$mean,nx=300,ny=200,asp=1, nlevel=80, main = "Mean of Prediction",axes=FALSE)
quilt.plot(grid_ex$X,grid_ex$Y, grid_ex$mean,nx=300,ny=200,asp=1, nlevel=80, main = "Mean of Prediction",axes=FALSE)
quilt.plot(grid_sp$X,grid_sp$Y, grid_sp$mean,nx=300,ny=200,asp=1, nlevel=80, main = "Mean of Prediction",axes=FALSE)


# same for standard deviation
quilt.plot(grid_lm$X,grid_lm$Y, grid_lm$sd,nx=300,ny=200,asp=1, nlevel=80, main = "Standard Error of Prediction",axes=FALSE)
quilt.plot(grid_me$X,grid_me$Y, grid_me$sd,nx=300,ny=200,asp=1, nlevel=80, main = "Standard Error of Prediction",axes=FALSE)
quilt.plot(grid_ex$X,grid_ex$Y, grid_ex$sd,nx=300,ny=200,asp=1, nlevel=80, main = "Standard Error of Prediction",axes=FALSE)
quilt.plot(grid_sp$X,grid_sp$Y, grid_sp$sd,nx=300,ny=200,asp=1, nlevel=80, main = "Standard Error of Prediction",axes=FALSE)

# Combine maps with PowerPoint into single picture
# Add additional infos there
```

### Combined map using ggplot

```{r comb_map}
#combine data
grid_ex$mean_lm <- grid_lm$mean
grid_ex$sd_lm <- grid_lm$sd
grid_ex$mean_sp <- grid_sp$mean
grid_ex$sd_sp <- grid_sp$sd
grid_ex$mean_me <- grid_me$mean
grid_ex$sd_me <- grid_me$sd

# prepare to go spatial
grid <- grid_ex
grid$ID <- 1:nrow(grid)
grid$X <- 1000 * grid$X
grid$Y <- 1000 * grid$Y




```

### Fitted versus measured

```{r plot_fitvsmeas}
aero_lm <- readRDS("../_15_noflight/Model_LM_inla/Output/aero_results.rds")
aero_me <- readRDS("../_15_noflight/Model_standard/Output/aero_results.rds")
aero_ex <- readRDS("../_15_noflight/Model_extended_priors/Main/Output/aero_results.rds")

#plot only for extended
ggplot(aero_ex, aes(x=exp(mean), y=TERR)) + geom_point(col = "black", alpha=0.1) + geom_abline(slope=1, intercept = 0,col="blue",size=1.2) + coord_fixed(ratio=1) + theme_minimal()+xlab("Predicted mean")+ylab("Measured")+ggtitle("Predicted vs. Measured")

# save
predvsmeas <- ggplot(aero_ex, aes(x=exp(mean), y=TERR)) + geom_point(col = "black", alpha=0.1) + geom_abline(slope=1, intercept = 0,col="blue",size=1.2) + coord_fixed(ratio=1) + theme_minimal()+xlab("Predicted mean [nSv/h]")+ylab("Measured [nSv/h]")
ggsave(predvsmeas, filename = "pred_vs_meas.jpg", dpi = "print")

```

### Fitted versus measured - normalized

```{r fitt_vs_meas2}
# how well is measured in predicted intervall?
# i will do: (LogTerr - mean.pred)/sd.pred
# ideally, this should follow a standard normal
aero_ex$norm <- (aero_ex$LogTerr - aero_ex$mean)/aero_ex$sd
plot(density(aero_ex$norm))

# I plot it against the predicted value and add ablines for the 95% intervall at 1.96
ggplot(aero_ex, aes(x=exp(mean), y=norm)) + geom_density_2d() +
  geom_abline(slope=0, intercept = 0,col="blue",size=1.2) +
  geom_abline(slope=0, intercept = -1.96,col="red",size=1) +
  geom_abline(slope=0, intercept = 1.96,col="red",size=1) +
  coord_fixed(ratio=1) + theme_minimal()+xlab("Predicted mean")+ylab("Residual/Sd of Prediction")+ggtitle("Density Plot predictions vs. normalized residuals")

# would be interesting to see density along x axis-- not sure how to best do...
# (sort of coverage probability by measured levels..)
```



### Maps of covariates

```{r covariate maps}
grid_ex$Preds <- exp(grid_ex$mean)
quilt.plot(grid_ex$X,grid_ex$Y, ifelse(grid_ex$Preds > 40,grid_ex$Preds,40),nx=300,ny=200,asp=1, nlevel=80, main = "Mean of Prediction",axes=FALSE) #screen values lower than 40nSv/h to enhance contrast?


grid_ex$Y <- grid_ex$Y*1000
grid_ex$X <- grid_ex$X*1000
coordinates(grid_ex) <- ~X+Y
proj4string(grid_ex) <- proj4string(CH_canton)
gridded(grid_ex) <-TRUE

tm_shape(CH_canton) + tm_borders(col = "slategrey") +
  tm_shape(CH_lakes) + tm_fill(col = "cadetblue1",alpha=0.4)+
  tm_shape(grid_ex) + tm_raster(col = "Preds", palette = "-plasma" ,style="cont",title = "Terrestrial Radiation [nSv/h]")+
  tm_layout(frame=FALSE, legend.text.size = 1.2, legend.title.size = 2)

terrmap <- tm_shape(CH_canton) + tm_borders(col = "slategrey") +
  tm_shape(CH_lakes) + tm_fill(col = "cadetblue1",alpha=0.4)+
  tm_shape(grid_ex) + tm_raster(col = "Preds", palette = "-plasma" ,style="cont",title = "Terrestrial Radiation [nSv/h]")+
  tm_layout(frame=FALSE, legend.text.size = 1.2, legend.title.size = 2 ,bg.color = "transparent") 
tmap_save(terrmap, "map_terrRad.jpg")

# Lithology
tm_shape(grid_ex) + tm_raster(col = "GESTEINKL", title = "Lithology") +
  tm_layout(legend.outside = TRUE, frame = FALSE, legend.text.size = 0.86, legend.title.size = 1.3)
Litho <- tm_shape(grid_ex) + tm_raster(col = "GESTEINKL", title = "Lithology") +
  tm_layout(legend.outside = TRUE, frame = FALSE, legend.text.size = 0.86, legend.title.size = 1.3)
tmap_save(Litho, "lithology.jpg")

# Tectonics
tm_shape(grid_ex) + tm_raster(col = "LEG_TEK_1", title = "Tectonic units") +
  tm_layout(legend.outside = TRUE, frame = FALSE, legend.title.size = 1.3)
Tecto <- tm_shape(grid_ex) + tm_raster(col = "LEG_TEK_1", title = "Tectonic units") +
  tm_layout(legend.outside = TRUE, frame = FALSE, legend.title.size = 1.3)
tmap_save(Tecto, "tectonics.jpg")

# Land cover
tm_shape(grid_ex) + tm_raster(col = "LC09_6", title = "Land coverage") +
  tm_layout(legend.outside = TRUE, frame = FALSE, legend.title.size = 1.3, legend.text.size = 1.2)
LC <- tm_shape(grid_ex) + tm_raster(col = "LC09_6", title = "Land coverage") +
  tm_layout(legend.outside = TRUE, frame = FALSE, legend.title.size = 1.3, legend.text.size = 1.2)
tmap_save(LC, "landcover.jpg")

# rain fall
ScaleRain <- readRDS("../_15_noflight/Input/SCALE_rain.rds")
grid_ex$rain <- grid_ex$rain*ScaleRain
tm_shape(grid_ex) + tm_raster(col = "rain", title = "Cumulative rainfall [mm]", style = "cont") +
  tm_layout(legend.outside = TRUE, frame = FALSE, legend.title.size = 1.5, legend.text.size = 1.1)
rainfall <- tm_shape(grid_ex) + tm_raster(col = "rain", title = "Cumulative rainfall [mm]", style = "cont") +
  tm_layout(legend.outside = TRUE, frame = FALSE, legend.title.size = 1.5, legend.text.size = 1.1)
tmap_save(rainfall, "rainfall.jpg")




```

### Variogram - Directional variogram

```{r plot_vrgs}
aero_cleaned$rain <- (aero_cleaned$Rain_april30 + aero_cleaned$Rain_may1 + aero_cleaned$Rain_may2 + aero_cleaned$Rain_may3 + aero_cleaned$Rain_may4 + aero_cleaned$Rain_may5)/1000
vrg <- variogram(TERR ~ GESTEINKL + LEG_TEK_1 + LC09_6 + rain, data = aero_cleaned[sample(1:nrow(aero_cleaned),100000),], 
                 boundaries = c(500,1000,1500,2000,2500,3000,4000,5000,6000,7000,8000,10000,12000,15000,18000,21000,24000,27000,30000,35000))
plot(vrg, xlab = "Distance [m]", ylab = expression(paste('Semivariance [(nSv/h) '^2,' ]',sep='')))
```

### Mesh u_1

Which mesh to do? I use the coarse mesh used to fit the extended model as example.

```{r exmaplemesh}
mesh <- readRDS("../_15_noflight/Model_extended_priors/Output/mesh_u1.rds")
plot(mesh)
# adapt title etc in PowerPoint
plot(mesh)
points(aero_ex$X,aero_ex$Y, col = "red")
```

### Distribution Measurements - Predictions

Compare distribution of measurements versus distribution of values predicted on map.

```{r distribution}
grid_ex <- readRDS("../_15_noflight/Model_extended_priors/Main/Output/grid_results.rds")

plot(density(exp(grid_ex$mean)), col = "black", lwd=2, xlab="Terrestrial Radiation [nSv/h]", main = "", bty = "n", cex.lab  = 1.4)
grid()
```


## Tables

### Fitted hyperparameters

see compare iter and extract_results scripts.

### Performances

see extract results and comereiter scripts for numbers.

### Fitted hyperparameters Sensitivity Analysis

See file sensitivity analysis.

### Performances Sensitivity Analysis

See file sensitivity analysis.

