---
title: "TablesForArticle"
author: "Folly Christophe"
date: "19 6 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages, results=F, warning=FALSE, message=FALSE, include=FALSE}
library(INLA)
library(Metrics)
library(fields)
library(tidyverse)
library(ggplot2)
library(scoringRules)
library(knitr)
library(gstat)
library(sp)
library(MatrixModels)
library(kableExtra)
```

## Introduction

Extract and display numbers and tables to be included in the article on the map.

## Tables

### Hyperpars

```{r hyperpars_extract}
#to be done
#init dataframe
df <- c("Model" = NA,"sd_epsilon" = NA, "range" = NA, "sd_field" = NA)
df <- data.frame(t(df))
#sd epsilon lm
res_lm <- readRDS("../Model_LM_inla/Output/results.rds")
df[1,] <- c("lm_inla", res_lm$summary.hyperpar[,1],NA,NA)
df <- rbind(df, c("lm_inla_sd", res_lm$summary.hyperpar[,2],NA,NA))
#pure spatial
res_sp <- readRDS("../Model_SPDE/Output/results.rds")
df <- rbind(df,c("pure_spatial", res_sp$summary.hyperpar[1,1], res_sp$summary.hyperpar[2,1], res_sp$summary.hyperpar[3,1]))
df <- rbind(df,c("pure_spatial_sd", res_sp$summary.hyperpar[1,2], res_sp$summary.hyperpar[2,2], res_sp$summary.hyperpar[3,2]))
#mixed
res_me <- readRDS("../Model_standard/Output/results.rds")
df <- rbind(df,c("mixed_effect", res_me$summary.hyperpar[1,1], res_me$summary.hyperpar[2,1], res_me$summary.hyperpar[3,1]))
df <- rbind(df,c("mixed_effect_sd", res_me$summary.hyperpar[1,2], res_me$summary.hyperpar[2,2], res_me$summary.hyperpar[3,2]))
#extended lm
res_eme <- readRDS("../Model_extended_priors/Main/Output/results.rds")
res_eme$summary.hyperpar[,1:2]
```

```{r hyperpars_display}
#do table of hyperpars
kable(df)
```

### Coefficients

```{r coeffs}
df2 <- as.data.frame(cbind(rownames(res_lm$summary.fixed) ,res_lm$summary.fixed[,1:2], res_me$summary.fixed[,1:2], res_eme$summary.fixed[,1:2]))
names(df2) <- c("Cov", "lm_mean", "lm_sd", "me_mean", "me_sd", "eme_mean", "eme_sd")
df2
```


### CV scores

I first create a dataframe to store the performance results.

```{r init_df, results=F}
performances <- data.frame(t(rep(NA,9)))
names(performances) <- c("model", "RMSEtrain","RMSErCV","RMSEspCV","R2train","R2rCV","R2spCV","CRPSrCV","CRPSspCV")
```

#### Linear Model (inla)


```{r lminla_load}
grid <- readRDS("../Model_LM_inla/Output/grid_results.rds")
fit <- readRDS("../Model_LM_inla/Output/aero_results.rds")
validr <- readRDS("../Model_LM_inla/CVr/Output/valid_results_cvr.rds")
valid1 <- readRDS("../Model_LM_inla/CV1/Output/valid_results_cv1.rds")
valid2 <- readRDS("../Model_LM_inla/CV2/Output/valid_results_cv2.rds")
valid3 <- readRDS("../Model_LM_inla/CV3/Output/valid_results_cv3.rds")
valid4 <- readRDS("../Model_LM_inla/CV4/Output/valid_results_cv4.rds")

#performance
RMSEtrain <- rmse(actual = fit$TERR, predicted = exp(fit$mean))
RMSErCV <- rmse(actual = validr$TERR, predicted = exp(validr$mean))
RMSEspCV <- rmse(actual = c(valid1$TERR, valid2$TERR[!is.na(valid2$mean)], valid3$TERR, valid4$TERR),
                 predicted = exp(c(valid1$mean, valid2$mean[!is.na(valid2$mean)], valid3$mean, valid4$mean)))
R2train <- cor(fit$TERR, exp(fit$mean))^2
R2rCV <- cor(validr$TERR, exp(validr$mean))^2
R2spCV <- cor(c(valid1$TERR, valid2$TERR[!is.na(valid2$mean)], valid3$TERR, valid4$TERR),
              exp(c(valid1$mean, valid2$mean[!is.na(valid2$mean)], valid3$mean, valid4$mean)))^2
valid <- rbind(valid1,valid2,valid3,valid4)
CRPSspCV <- mean(crps(y = valid$LogTerr, family = "normal", mean = valid$mean, sd = valid$sd)) #on log-scale
CRPSrCV <- mean(crps(y = validr$LogTerr, family = "normal", mean = validr$mean, sd = validr$sd))

#add
performances[1,] <- c("lM_inla", RMSEtrain,RMSErCV,RMSEspCV,R2train,R2rCV,R2spCV,CRPSrCV, CRPSspCV)
performances[1,]
```


#### pure correlated spatial random-effects

```{r pure_sp_load}
grid <- readRDS("../Model_SPDE/Output/grid_results.rds")
fit <- readRDS("../Model_SPDE/Output/aero_results.rds")
validr <- readRDS("../Model_SPDE/CVr/Output/valid_results_cvr.rds")
valid1 <- readRDS("../Model_SPDE/CV1/Output/valid_results_cv1.rds")
valid2 <- readRDS("../Model_SPDE/CV2/Output/valid_results_cv2.rds")
valid3 <- readRDS("../Model_SPDE/CV3/Output/valid_results_cv3.rds")
valid4 <- readRDS("../Model_SPDE/CV4/Output/valid_results_cv4.rds")

#compute perf
RMSEtrain <- rmse(actual = fit$TERR, predicted = exp(fit$mean))
RMSErCV <- rmse(actual = validr$TERR, predicted = exp(validr$mean))
RMSEspCV <- rmse(actual = c(valid1$TERR, valid2$TERR[!is.na(valid2$mean)], valid3$TERR, valid4$TERR),
                 predicted = exp(c(valid1$mean, valid2$mean[!is.na(valid2$mean)], valid3$mean, valid4$mean)))
R2train <- cor(fit$TERR, exp(fit$mean))^2
R2rCV <- cor(validr$TERR, exp(validr$mean))^2
R2spCV <- cor(c(valid1$TERR, valid2$TERR[!is.na(valid2$mean)], valid3$TERR, valid4$TERR),
              exp(c(valid1$mean, valid2$mean[!is.na(valid2$mean)], valid3$mean, valid4$mean)))^2
valid <- rbind(valid1,valid2,valid3,valid4)
CRPSspCV <- mean(crps(y = valid$LogTerr, family = "normal", mean = valid$mean, sd = valid$sd)) #on log-scale
CRPSrCV <- mean(crps(y = validr$LogTerr, family = "normal", mean = validr$mean, sd = validr$sd))

#add
performances[2,] <- c("pure_spatial", RMSEtrain,RMSErCV,RMSEspCV,R2train,R2rCV,R2spCV,CRPSrCV, CRPSspCV)
performances[2,]

rm(grid,validr,valid1,valid2,valid3,valid4,fit,RMSEtrain,RMSErCV,RMSEspCV,R2train,R2rCV,R2spCV,CRPSspCV)
```

### mixed-effects model

```{r mixedeffects_load}
grid <- readRDS("../Model_standard/Output/grid_results.rds")
fit <- readRDS("../Model_standard/Output/aero_results.rds")
validr <- readRDS("../Model_standard/CVr/Output/valid_results_cvr.rds")
valid1 <- readRDS("../Model_standard/CV1/Output/valid_results_cv1.rds")
valid2 <- readRDS("../Model_standard/CV2/Output/valid_results_cv2.rds")
valid3 <- readRDS("../Model_standard/CV3/Output/valid_results_cv3.rds")
valid4 <- readRDS("../Model_standard/CV4/Output/valid_results_cv4.rds")

#comp perf
RMSEtrain <- rmse(actual = fit$TERR, predicted = exp(fit$mean))
RMSErCV <- rmse(actual = validr$TERR, predicted = exp(validr$mean))
RMSEspCV <- rmse(actual = c(valid1$TERR, valid2$TERR[!is.na(valid2$mean)], valid3$TERR, valid4$TERR),
                 predicted = exp(c(valid1$mean, valid2$mean[!is.na(valid2$mean)], valid3$mean, valid4$mean)))
R2train <- cor(fit$TERR, exp(fit$mean))^2
R2rCV <- cor(validr$TERR, exp(validr$mean))^2
R2spCV <- cor(c(valid1$TERR, valid2$TERR[!is.na(valid2$mean)], valid3$TERR, valid4$TERR),
              exp(c(valid1$mean, valid2$mean[!is.na(valid2$mean)], valid3$mean, valid4$mean)))^2
valid <- rbind(valid1,valid2,valid3,valid4)
CRPSspCV <- mean(crps(y = valid$LogTerr, family = "normal", mean = valid$mean, sd = valid$sd)) #on log-scale
CRPSrCV <- mean(crps(y = validr$LogTerr, family = "normal", mean = validr$mean, sd = validr$sd))

#add
performances[3,] <- c("mixed-effects", RMSEtrain,RMSErCV,RMSEspCV,R2train,R2rCV,R2spCV,CRPSrCV, CRPSspCV)
performances[3,]

rm(grid,validr,valid1,valid2,valid3,valid4,fit,RMSEtrain,RMSErCV,RMSEspCV,R2train,R2rCV,R2spCV,CRPSspCV)
```

#### extended mixed effects

```{r perf_ext_me}
#load informative prior
grid <- readRDS("../Model_extended_priors/Main/Output/grid_results.rds")
fit <- readRDS("../Model_extended_priors/Main/Output/aero_results.rds")
validr <- readRDS("../Model_extended_priors/CVr/Output/valid_results_cvr.rds") 
valid1 <- readRDS("../Model_extended_priors/CV1/Output/valid_results_cv1.rds")
valid2 <- readRDS("../Model_extended_priors/CV2/Output/valid_results_cv2.rds")
valid3 <- readRDS("../Model_extended_priors/CV3/Output/valid_results_cv3.rds")
valid4 <- readRDS("../Model_extended_priors/CV4/Output/valid_results_cv4.rds")
valid <- rbind(valid1,valid2,valid3,valid4)
valid$pred <- valid$mean
valid$sd <- valid$sd
validr$pred <- validr$mean
validr$sd <- validr$sd

# performance iterative
RMSEtrain <- rmse(actual = fit$TERR, predicted = exp(fit$mean))
RMSErCV <- rmse(actual = validr$TERR, predicted = exp(validr$pred))
RMSEspCV <- rmse(actual = valid$TERR,
                 predicted = exp(valid$pred))
R2train <- cor(fit$TERR, exp(fit$mean))^2
R2rCV <- cor(validr$TERR, exp(validr$pred))^2
R2spCV <- cor(valid$TERR,exp(valid$pred))^2
CRPSspCV <- mean(crps(y = valid$LogTerr, family = "normal", mean = valid$pred, sd = valid$sd)) #on log-scale 
CRPSrCV <- mean(crps(y = validr$LogTerr, family = "normal", mean = validr$mean, sd = validr$sd))

#add
#add
performances[4,] <- c("extended mixed-effects", RMSEtrain,RMSErCV,RMSEspCV,R2train,R2rCV,R2spCV,CRPSrCV, CRPSspCV)
performances[4,]

```

```{r performances}
performances
```

## Additional numbers

```{r clean_ws1}
rm(list=ls())
```

### Number of mesh nodes sensitivity 

```{r meshnodes}
#extended
readRDS("../Model_extended_priors/Output/mesh_u1.rds")$n
readRDS("../Model_extended_priors/Output/mesh_u2.rds")$n
#sensitivity denser
readRDS("../Model_extended_priors_sensitivity_meshsmall/Output/mesh_u1.rds")$n
readRDS("../Model_extended_priors_sensitivity_meshsmall/Output/mesh_u2.rds")$n
#sensitivity less dense
readRDS("../Model_extended_priors_sensitivity_meshlarge/Output/mesh_u1.rds")$n
readRDS("../Model_extended_priors_sensitivity_meshlarge/Output/mesh_u2.rds")$n
```

### Overview Number of Measurements by category (total and included)

```{r measurements by category}
# ====  measurements included
data <- readRDS("../Model_extended_priors/Input/aero_cluster.rds")

#litho
data %>% group_by(GESTEINKL) %>% summarise("N" = n(), "mean" = mean(TERR), "min" = quantile(TERR, probs = c(0)), "25q" = quantile(TERR, probs = c(0.25)), "median" = quantile(TERR, probs = c(0.5)), "75q" = quantile(TERR, probs = c(0.75)), max = quantile(TERR, probs = c(1)))

#tecto
data %>% group_by(LEG_TEK_1) %>% summarise("N" = n(), "mean" = mean(TERR), "min" = quantile(TERR, probs = c(0)), "25q" = quantile(TERR, probs = c(0.25)), "median" = quantile(TERR, probs = c(0.5)), "75q" = quantile(TERR, probs = c(0.75)), max = quantile(TERR, probs = c(1)))

#lc
data %>% group_by(LC09_6) %>% summarise("N" = n(), "mean" = mean(TERR), "min" = quantile(TERR, probs = c(0)), "25q" = quantile(TERR, probs = c(0.25)), "median" = quantile(TERR, probs = c(0.5)), "75q" = quantile(TERR, probs = c(0.75)), max = quantile(TERR, probs = c(1)))



# ====  excluded data
library(sf)
data <- st_set_geometry(readRDS("../Model_extended_priors/Input/excluded_measurements.rds"),NULL)

#litho
data %>% group_by(GESTEINKL) %>% summarise("N" = n(), "mean" = mean(TERR), "min" = quantile(TERR, probs = c(0)), "25q" = quantile(TERR, probs = c(0.25)), "median" = quantile(TERR, probs = c(0.5)), "75q" = quantile(TERR, probs = c(0.75)), max = quantile(TERR, probs = c(1)))

#tecto
data %>% group_by(LEG_TEK_1) %>% summarise("N" = n(), "mean" = mean(TERR), "min" = quantile(TERR, probs = c(0)), "25q" = quantile(TERR, probs = c(0.25)), "median" = quantile(TERR, probs = c(0.5)), "75q" = quantile(TERR, probs = c(0.75)), max = quantile(TERR, probs = c(1)))

#lc
data %>% group_by(LC09_6) %>% summarise("N" = n(), "mean" = mean(TERR), "min" = quantile(TERR, probs = c(0)), "25q" = quantile(TERR, probs = c(0.25)), "median" = quantile(TERR, probs = c(0.5)), "75q" = quantile(TERR, probs = c(0.75)), max = quantile(TERR, probs = c(1)))


# ==== all afgter cleaning
data <- rbind(st_set_geometry(readRDS("../Model_extended_priors/Input/excluded_measurements.rds"),NULL)[,c("TERR","GESTEINKL","LEG_TEK_1","LC09_6")],readRDS("../Model_extended_priors/Input/aero_cluster.rds")[,c("TERR","GESTEINKL","LEG_TEK_1","LC09_6")])

#litho
data %>% group_by(GESTEINKL) %>% summarise("N" = n(), "mean" = mean(TERR), "min" = quantile(TERR, probs = c(0)), "25q" = quantile(TERR, probs = c(0.25)), "median" = quantile(TERR, probs = c(0.5)), "75q" = quantile(TERR, probs = c(0.75)), max = quantile(TERR, probs = c(1)))

#tecto
data %>% group_by(LEG_TEK_1) %>% summarise("N" = n(), "mean" = mean(TERR), "min" = quantile(TERR, probs = c(0)), "25q" = quantile(TERR, probs = c(0.25)), "median" = quantile(TERR, probs = c(0.5)), "75q" = quantile(TERR, probs = c(0.75)), max = quantile(TERR, probs = c(1)))

#lc
data %>% group_by(LC09_6) %>% summarise("N" = n(), "mean" = mean(TERR), "min" = quantile(TERR, probs = c(0)), "25q" = quantile(TERR, probs = c(0.25)), "median" = quantile(TERR, probs = c(0.5)), "75q" = quantile(TERR, probs = c(0.75)), max = quantile(TERR, probs = c(1)))
```





















